### 2023-02-22 code/translocation_survival_analysis.Rmd
Change plot order of conditional effects figure
Edit figures, add those and other figures to manuscript

mcmc_intervals(m1a, point_est = "median", prob = 0.80, prob_outer = 0.9) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  # add 0-line to accentuate default line
  ggtitle("Model with non-standardized predictors only",
          "Posterior distributions with medians & 90% uncertainty intervals")

mcmc_areas(m1a, area_method = "equal height", prob = 0.90) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Model with non-standardized predictors only",
          "Posterior distributions with medians & 90% uncertainty intervals")

@@

-   Using non-standardized predictors makes plot hard to read due to vastly different scales of predictors.\
-   Based on 90% uncertainty intervals, important predictors are snow_t1, elevation, first, donor70567, and donor72996.

@@

mcmc_areas(m1b, area_method = "equal height", prob = 0.90) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Model with standardized predictors only",
          "Posterior distributions with medians & 90% uncertainty intervals")

@@

-   As expected, using standardized predictors allows easy comparison across predictors. Important predictors are unchanged.
-   Based on 90% uncertainty intervals, important predictors are snow_t1, elevation, first, donor70567, and donor72996.

@@

mcmc_areas(m1c, area_method = "equal height", prob = 0.90, pars = b1) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Model with standardized predictors & site_id as grouping variable",
          "Posterior distributions with medians & 90% uncertainty intervals")

@@

-   Based on 90% uncertainty intervals, important predictors reduced to elevation and male.

@@

-   Based on loo, model with translocation_id as grouping variable provides slightly better fit than model with site_id as grouping variable.
-   Based on 90% uncertainty intervals, important predictors are similar to those in model without grouping variable: snowt1_z, elev_z, male_c0.548, donor72996.

@@

mcmc_areas(m1e, area_method = "equal height", prob = 0.90, pars = b1) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Model with standardized predictors & translocation_id/site_id as grouping variables",
          "Posterior distributions with medians & 90% uncertainty intervals")

@@

p_sex <- d1_malec %>% 
  select(site_id, translocation_id, snowt1_z, elev_z, male_c, sex, donor, surv) %>% 
  add_epred_draws(m1d_reduce, ndraws = 100, re_formula = NA) %>% 
  ggplot(aes(x = male, y = surv)) +
    geom_boxplot(aes(y = .epred, x = sex), fill = "lightgray") +
    labs(x = "Sex", y = "P(1-year frog survival)") +
    annotate(geom = "text", x = 0.4, y = 0.94, hjust = -1, label = "C", size = 4) +
    theme_classic()

@@

p_donor <- d1_donor %>% 
  select(site_id, translocation_id, snowt1_z, elev_z, male_c, donor, surv) %>% 
  add_epred_draws(m1d_reduce, ndraws = 100, re_formula = NA) %>% 
  ggplot(aes(x = donor, y = surv)) +
    geom_boxplot(aes(y = .epred, x = donor), fill = "lightgray") +
    labs(x = "Donor population", y = "P(1-year frog survival)") +
    annotate(geom = "text", x = 3.3, y = 0.83, hjust = 0, label = "D", size = 4) +
    theme_classic()

@@

```{r}
cond_effects_plot <- (p_snowt1 | p_elev) / (p_sex | p_donor)
```

@@

mcmc_areas(m2a, area_method = "equal height", prob = 0.90, pars = b2) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Model with unstandardized predictor & no grouping variable",
          "Posterior distributions with medians & 90% uncertainty intervals")

@@

## Model with standardized predictor, no grouping variable

@@

mcmc_areas(m2b, area_method = "equal height", prob = 0.90, pars = b3) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Model with standardized predictor & no grouping variable",
          "Posterior distributions with medians & 90% uncertainty intervals")

@@

## Model with standardized predictor, translocation_id grouping variable

@@

mcmc_areas(m2c, area_method = "equal height", prob = 0.90, pars = b3) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Model with standardized predictor & translocation_id as grouping variable",
          "Posterior distributions with medians & 90% uncertainty intervals")

@@

## Model with standardized predictor, site_id grouping variable

@@

mcmc_areas(m2d, area_method = "equal height", prob = 0.90, pars = b3) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Model with standardized predictor & site_id as grouping variable",
          "Posterior distributions with medians & 90% uncertainty intervals")

>>>>

mcmc_intervals(m1a, point_est = "median", prob = 0.80, prob_outer = 0.9) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  # add 0-line to accentuate default line
  ggtitle("Model with non-standardized predictors only",
          "Posterior distributions with medians & 95% uncertainty intervals")

mcmc_areas(m1a, area_method = "equal height", prob = 0.95) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Model with non-standardized predictors only",
          "Posterior distributions with medians & 95% uncertainty intervals")

@@

-   Using non-standardized predictors makes plot hard to read due to vastly different scales of predictors.\
-   Based on 95% uncertainty intervals, important predictors are snow_t1, elevation, first, donor70567, and donor72996.

@@

mcmc_areas(m1b, area_method = "equal height", prob = 0.95) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Model with standardized predictors only",
          "Posterior distributions with medians & 95% uncertainty intervals")

@@

-   As expected, using standardized predictors allows easy comparison across predictors. Important predictors are unchanged.
-   Based on 95% uncertainty intervals, important predictors are snow_t1, elevation, first, donor70567, and donor72996.

@@

mcmc_areas(m1c, area_method = "equal height", prob = 0.95, pars = b1) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Model with standardized predictors & site_id as grouping variable",
          "Posterior distributions with medians & 95% uncertainty intervals")

@@

-   Based on 95% uncertainty intervals, important predictors reduced to elevation and male.

@@

-   Based on loo, model with translocation_id as grouping variable provides slightly better fit than model with site_id as grouping variable.
-   Based on 95% uncertainty intervals, important predictors are similar to those in model without grouping variable: snowt1_z, elev_z, male_c0.548, donor72996.

@@

mcmc_areas(m1e, area_method = "equal height", prob = 0.95, pars = b1) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Model with standardized predictors & translocation_id/site_id as grouping variables",
          "Posterior distributions with medians & 95% uncertainty intervals")

@@

p_sex <- d1_malec %>% 
  select(site_id, translocation_id, snowt1_z, elev_z, male_c, sex, donor, surv) %>% 
  add_epred_draws(m1d_reduce, ndraws = 100, re_formula = NA) %>% 
  ggplot(aes(x = male, y = surv)) +
    geom_boxplot(aes(y = .epred, x = sex), fill = "lightgray") +
    labs(x = "Sex", y = "P(1-year frog survival)") +
    annotate(geom = "text", x = 0.4, y = 0.94, hjust = -1, label = "D", size = 4) +
    theme_classic()

@@

p_donor <- d1_donor %>% 
  select(site_id, translocation_id, snowt1_z, elev_z, male_c, donor, surv) %>% 
  add_epred_draws(m1d_reduce, ndraws = 100, re_formula = NA) %>% 
  ggplot(aes(x = donor, y = surv)) +
    geom_boxplot(aes(y = .epred, x = donor), fill = "lightgray") +
    labs(x = "Donor population", y = "P(1-year frog survival)") +
    annotate(geom = "text", x = 3.3, y = 0.94, hjust = 0, label = "C", size = 4) +
    theme_classic()

@@

```{r}
cond_effects_plot <- (p_snowt1 | p_elev) / (p_donor | p_sex)
```

@@

mcmc_areas(m2a, area_method = "equal height", prob = 0.95, pars = b2) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Model with unstandardized predictor & no grouping variable",
          "Posterior distributions with medians & 95% uncertainty intervals")

@@

## Model with standardized predictors, no grouping variable

@@

labels_alt <- c("survival_trans1", "(Intercept)")
mcmc_areas(m2b, area_method = "equal height", prob = 0.95, pars = b3) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +
  scale_y_discrete(labels = labels_alt, limits = rev)
#  ggtitle("Model with standardized predictor & no grouping variable",
#          "Posterior distributions with medians & 95% uncertainty intervals")

@@

### Save plot
```{r}
ggsave(here::here("doc", "manuscript", "figures", "mcmc_areas_m2b.png"), bg = "white", width = 7, height = 4)
```

## Model with standardized predictors, translocation_id grouping variable

@@

mcmc_areas(m2c, area_method = "equal height", prob = 0.95, pars = b3) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Model with standardized predictor & translocation_id as grouping variable",
          "Posterior distributions with medians & 95% uncertainty intervals")

@@

## Model with standardized predictors, site_id grouping variable

@@

mcmc_areas(m2d, area_method = "equal height", prob = 0.95, pars = b3) +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed", lwd = 0.5) +  
  ggtitle("Model with standardized predictor & site_id as grouping variable",
          "Posterior distributions with medians & 95% uncertainty intervals")

### 2023-02-22 doc/manuscript/translocation.qmd
Edit figures, add those and other figures to manuscript

Reintroductions of *R. sierrae* were conducted in Yosemite National Park (California, USA) as part of an effort to reestablish populations of this endangered species. These reintroductions involved the collection of adult frogs from large, Bd-positive donor populations <!--# Add figure to SI showing Bd loads before and after translocations -->that have recovered in recent years following Bd-caused declines [@knapp2016] and translocating them to nearby recipient sites. We refer to these reintroductions as "translocations". During the period `r min(d1$year)`--`r max(d1$year)`, we conducted `r n_distinct(d1$site_id, d1$date)` translocations of frogs to `r n_distinct(d1$site_id)` recipient sites (`r transno_min`--`r transno_max` translocations per site; @fig-yosemap) <!--# How to restart numbering for figures in SI, add "S" prefix? -->. Each translocated cohort included `r transfrogs_min` to `r transfrogs_max` frogs collected from one of `r n_distinct(d1$donor)` donor populations (@fig-yosemap). Estimates of frog survival during the one year following translocation indicate that survival was highly variable between sites, but relatively constant within sites (for the subset of sites that received multiple translocations; @fig-translocation-survival). As such, the average survival of frogs in the first translocation to a site was an important predictor of survival of frogs in all subsequent translocations to that site (median of posterior distribution = 4.14, uncertainty interval = 3.06\--5.28 <!--# How to extract these values from table produced by line 934 in translocation_survival_analysis.Rmd (or from associated code output)? --> ; see Supporting Information for analysis details), and suggests an important effect of site characteristics and/or (site \* frog) interactions on frog survival. For frogs translocated to each site, Bd load changed little between the periods immediately prior to translocation versus during the one year after translocation (see SI for analysis details).

To identify drivers of post-translocation frog survival, we developed a hierarchical model in a Bayesian framework (see SI for details). Our specific interest was whether Bd load had an important effect on adult survival, as would be expected if the effect of frog resistance was insufficient to overcome the negative effect of Bd. The best model of 1-year frog survival identified several important predictors, but Bd load at the time of translocation was not among them (bd_load; @fig-survival-postdens). Important predictors included winter severity in the year following translocation (snow_t1), site elevation (elevation), and donor population (donor_72996; @fig-cond-effects). Males had somewhat higher survival than females, but this effect was only marginally important (sex_male; @fig-cond-effects). In addition to Bd load, effects of frog size (length), winter severity in the year of translocation (snow_t), day of year on which a translocation was conducted (day), and translocation order (order_first) were all unimportant (@fig-survival-postdens). In summary, our analyses show that Bd load is not a predictor of the survival of translocated frogs, consistent with frogs in recovering populations having evolved increased resistance and/or tolerance to Bd infection. In addition, the small changes in Bd load following translocation and relatively high survival in the majority of translocated cohorts suggests that putatively resistant frogs from recovering populations can maintain the benefits of resistance (i.e., low-moderate Bde loads, minimal Bd-induced mortality despite ongoing infection) in non-natal habitats. Finally, in three locations where longer mark-recapture time series allowed us to assess the survival of adults recruited to the population that are the progeny of translocated individuals, naturally recruited adults had equivalent or higher survival probabilities than the originally translocated adults <!--# Show results. -->. Collectively, these results show that putative resistance in non-natal habitats is maintained across generations of frogs.

@@

![Conditional effects of the important predictors of 1-year frog survival: (A) elevation of recipient site, (B) winter severity in the year following translocation, (C) sex, and (D) donor population. In (A) and (B), blue lines are medians and gray ribbons are 95% uncertainty intervals. In (C) and (D), box plots show medians, first and third quartiles, largest and smallest values within 1.5x interquartile range, and values outside the 1.5x interquartile range. <!--# Switch C and D panels. -->](figures/cond_effects_plot.png){#fig-cond-effects}

@@

### Materials and methods

@@

@@

![For the best model identifying predictors of 1-year frog survival following translocation, estimated posterior density curves and shaded 95% uncertainty intervals for all predictor variables. Variables are considered important predictors if the associated uncertainty interval does not overlap zero (dashed red line).](figures/mcmc_areas_m1d.png){#fig-survival-postdens width="450"}

![For frogs translocated to each recipient site, change in Bd load between the periods immediately prior to translocation versus during the 1-year period after translocation.](figures/bdload_beforeafter.png){#fig-bdload-beforeafter width="450"}

>>>>

Reintroductions of *R. sierrae* were conducted in Yosemite National Park (California, USA) as part of an effort to reestablish populations of this endangered species. These reintroductions involved the collection of adult frogs from large, Bd-positive donor populations that have recovered in recent years following Bd-caused declines [@knapp2016] and translocating them to nearby recipient sites. We refer to these reintroductions as "translocations". During the period `r min(d1$year)`--`r max(d1$year)`, we conducted `r n_distinct(d1$site_id, d1$date)` translocations of frogs to `r n_distinct(d1$site_id)` recipient sites (`r transno_min`--`r transno_max` translocations per site; @fig-yosemap) <!--# How to restart numbering for figures in SI, add "S" prefix? -->. Each translocated cohort included `r transfrogs_min` to `r transfrogs_max` frogs collected from one of `r n_distinct(d1$donor)` donor populations (@fig-yosemap). Estimates of frog survival during the one year following translocation indicate that survival was highly variable between recipient sites, but relatively constant within recipient sites (for the subset of sites that received multiple translocations; @fig-translocation-survival). As such, the average survival of frogs in the first translocation to a site was an important predictor of survival of frogs in all subsequent translocations to that site (@fig-transsurvival-postdens; see **Materials and Methods - Frog translocations** for analysis details), and suggests an important effect of site characteristics and/or (site \* frog) interactions on frog survival. For translocated frogs, Bd load changed little between immediately prior to translocation versus during the one year after translocation (@fig-bdload-beforeafter; see SI for analysis details). <!--# Add text describing that loads were below lethal threshold -->

To identify drivers of post-translocation frog survival, we developed a multilevel model in a Bayesian framework (see **Materials and Methods - Frog translocations** for analysis details). Our specific interest was whether Bd load had an important effect on adult survival, as would be expected if the effect of frog resistance was insufficient to overcome the negative effect of Bd. The best model of 1-year frog survival identified several important predictors, but Bd load at the time of translocation was not among them (bd_load; @fig-survival-postdens). Important predictors included winter severity in the year following translocation (snow_t1), site elevation (elevation), and donor population (donor_72996; @fig-cond-effects). Males had somewhat higher survival than females, but this effect was only marginally important (sex_male; @fig-cond-effects). In addition to Bd load, effects of frog size (length), winter severity in the year of translocation (snow_t), day of year on which a translocation was conducted (day), and translocation order (order_first) were all unimportant (@fig-survival-postdens). In summary, our analyses show that Bd load is not a predictor of the survival of translocated frogs, consistent with frogs in recovering populations having evolved sufficient resistance and/or tolerance to minimize the adverse effects of Bd infection. In addition, the small changes in Bd load following translocation and relatively high survival of frogds in the majority of translocated cohorts suggests that resistant frogs from recovering populations can maintain the benefits of resistance (i.e., low-moderate Bd loads, minimal Bd-induced mortality despite ongoing infection) in non-natal habitats. Finally, in three locations where longer mark-recapture time series allowed us to assess the survival of adults recruited to the population that are the progeny of translocated individuals, naturally recruited adults had equivalent or higher survival probabilities than the originally translocated adults <!--# Show results. -->. Collectively, these results show that putative resistance in non-natal habitats is maintained across generations of frogs.

@@

![Conditional effects of the important predictors of 1-year frog survival: (A) elevation of recipient site, (B) winter severity in the year following translocation, (C) donor population, and (D) sex. In (A) and (B), blue lines are medians and gray ribbons are 95% uncertainty intervals. In (C) and (D), box plots show medians, first and third quartiles, largest and smallest values within 1.5x interquartile range, and values outside the 1.5x interquartile range.](figures/cond_effects_plot.png){#fig-cond-effects}

@@

### Materials and Methods

@@

<!--# Add text and results describing Bd load before versus after translocation. -->

@@

![For the best model identifying predictors of 1-year frog survival following translocation, estimated posterior density curves and shaded 95% uncertainty intervals for the intercept and all predictor variables. Variables are considered important predictors if the associated uncertainty interval does not overlap zero (dashed red line).](figures/mcmc_areas_m1d.png){#fig-survival-postdens width="450"}

![For frogs translocated to each recipient site, Bd loads for the period immediately prior to translocation versus during the 1-year period after translocation.](figures/bdload_beforeafter.png){#fig-bdload-beforeafter width="450"}

![For the best model describing the average 1-year survival of frogs in the first translocation to a site as a predictor of survival of frogs in all subsequent translocations to that site, estimated posterior density curves and shaded 95% uncertainty intervals for the intercept and single predictor variable. A variable is considered an important predictor if the associated uncertainty interval does not overlap zero (dashed red line).](figures/mcmc_areas_m2b.png){#fig-transsurvival-postdens width="450"}
